% ========================================================
% 3method.tex
% ========================================================
\section{Methodology}
\label{sec:method}

\subsection{Problem Formulation}

Consider a transit venue with $N$ wireless APs indexed by
$\mathcal{N} = \{1, \dots, N\}$. At each discrete timestep $t$ (5-minute
intervals), each AP $i$ chooses a hybrid action
$\mathbf{a}_i^t = (p_i, c_i, \mathbf{w}_i) \in \mathcal{A}$, where
$p_i \in [P_{\min}, P_{\max}]$ is transmit power in dBm,
$c_i \in \{1, \dots, C\}$ is the assigned channel (Gumbel-softmax
relaxation), and $\mathbf{w}_i \in \Delta^3$ is a 3-simplex bandwidth
allocation across QoS classes (VoIP, streaming, best-effort).

The objective is to jointly maximise a multi-dimensional reward:
\begin{equation}
  r_i^t = \alpha r_{\mathrm{tp}} + \beta r_{\mathrm{lat}} +
          \gamma r_{\mathrm{fair}} + \delta r_{\mathrm{qos}} +
          \delta_p \, b_{\mathrm{pred}}
  \label{eq:reward}
\end{equation}
subject to an adaptive capacity constraint $L_i^t \le 0.9$
enforced via Lagrangian multiplier $\lambda_i^t$.
Parameters $\alpha = 0.40$, $\beta = 0.25$, $\gamma = 0.20$,
$\delta = 0.05$, $\delta_p = 0.10$ are tuned per the design specification.

The global MDP is factorised into per-AP sub-MDPs sharing parameters
through hierarchical federated aggregation (Section~\ref{sec:hierfed}).

\subsection{Module 1: ST-GCAT Predictor}

\subsubsection{Spatial GAT Layer}
Let $\mathbf{A}^t \in \Rset^{N \times N}$ denote the interference graph at
time $t$, where edge weight $a_{ij}$ encodes the RSSI-based interference
between APs $i$ and $j$. The spatial GAT applies multi-head attention:
\begin{equation}
  \mathbf{h}_i^{\mathrm{sp}} =
    \sigma\!\left(\sum_{j \in \mathcal{N}_i}
      \alpha_{ij}^{\mathrm{GAT}} \mathbf{W}^{\mathrm{sp}} \mathbf{x}_j\right),
  \quad
  \alpha_{ij}^{\mathrm{GAT}} =
    \frac{\exp(\mathrm{LeakyReLU}(\mathbf{a}^{\top}
      [\mathbf{W}\mathbf{h}_i \| \mathbf{W}\mathbf{h}_j]))}
    {\sum_{k}\exp(\mathrm{LeakyReLU}(\mathbf{a}^\top
      [\mathbf{W}\mathbf{h}_i \| \mathbf{W}\mathbf{h}_k]))}
  \label{eq:gat}
\end{equation}

\subsubsection{Temporal Transformer with Schedule Cross-Attention}
The temporal Transformer processes the history sequence
$\mathbf{H} = [\mathbf{h}^{t-T}, \dots, \mathbf{h}^t] \in \Rset^{N \times T \times F}$
with standard multi-head self-attention. Crucially, the schedule feature
tensor $\mathbf{S}^{t:t+\Delta} \in \Rset^{N \times \Delta \times D_s}$
(encoding flight/train departure counts per AP zone, dwell time, and gate
assignments) serves as the cross-attention Key and Value:
\begin{align}
  \mathbf{Z}_i &= \mathrm{CrossAttn}(Q = \mathbf{H}_i, K = \mathbf{S}_i,
                                       V = \mathbf{S}_i), \label{eq:crossattn} \\
  \hat{L}_i^{t+1:t+\Delta} &= \mathrm{PredHead}(\mathbf{Z}_i).
  \label{eq:pred}
\end{align}
The fused embedding $\bfz_\mathrm{fused} = [\mathbf{Z}_1, \dots, \mathbf{Z}_N]
\in \Rset^{N \times \Delta \times d_h}$ is stored and passed to the SAC
agent. Prediction targets for the joint loss are the next-step true loads
$L^{t+1}$ (step 0) and steady-state estimates for steps $1:\Delta-1$.

\subsection{Module 2: Graph-SAC Agent}
\label{sec:sac}

\subsubsection{State Construction}
The state of AP $i$ at time $t$ is:
\begin{equation}
  \mathbf{s}_i^t = \left[
    \mathbf{h}_i,\;
    \mathrm{vec}(\mathbf{Z}_{\mathrm{fused},i}),\;
    \mathbf{q}_i,\;
    \mathbf{s}^{\mathrm{sched}}_i
  \right] \in \Rset^{d_s},
  \label{eq:state}
\end{equation}
where $\mathbf{h}_i \in \Rset^5$ is the local feature vector,
$\mathrm{vec}(\mathbf{Z}_{\mathrm{fused},i}) \in \Rset^{\Delta \cdot d_h}$
(= 768 dims with $\Delta=6$, $d_h=128$) is the flattened graph embedding,
$\mathbf{q}_i \in \Rset^3$ contains QoS tail fractions, and
$\mathbf{s}^{\mathrm{sched}}_i \in \Rset^8$ encodes upcoming schedule events.
Total state dimension: $d_s = 5 + 768 + 3 + 8 = 784$.

\subsubsection{Hybrid Actor}
The actor produces a hybrid action via three heads:
\begin{itemize}
  \item \textbf{Power}: Gaussian reparameterisation with sigmoid squashing.
  Log-prob correction uses the correct sigmoid Jacobian:
  $\log|\sigma'(u)| = \log(\sigma(u)(1 - \sigma(u)))$.
  \item \textbf{Channel}: Gumbel-softmax with temperature $\tau$ (annealed
  from 2.0 to 0.5 during training), enabling differentiable discrete channel
  selection.
  \item \textbf{Bandwidth}: Dirichlet distribution parametrised by
  $\text{softplus}(\cdot) + 1$ concentrations, ensuring the 3-simplex
  bandwidth vector sums to 1.
\end{itemize}

\subsubsection{Soft Actor-Critic Update}
Standard SAC updates for critic (double-Q, Eq.~\ref{eq:critic})
and actor (Eq.~\ref{eq:actor}) with entropy coefficient $\alpha$
automatically tuned via:
\begin{align}
  \mathcal{L}_Q &= \mathbb{E}\!\left[
    (Q(s,a) - r - \gamma \bar{Q}(s', \bar{a}') + \alpha \log \pi(\bar{a}'|s'))^2
  \right], \label{eq:critic} \\
  \mathcal{L}_\pi &= \mathbb{E}\!\left[
    \alpha \log \pi(a|s) - Q(s,a)
  \right]. \label{eq:actor}
\end{align}

\subsubsection{Bidirectional Joint Gradient (Core Novelty)}
\label{sec:joint}
After each SAC update cycle, the predictor parameters $\phi$ are updated by
a joint loss that couples control objectives into the representation:
\begin{equation}
  \mathcal{L}_\mathrm{pred} =
    \underbrace{\frac{1}{N\Delta}\sum_{i,k}\!\left(
      \hat{L}_{i,k} - L^*_{i,k}
    \right)^2}_{\text{MSE, all } \Delta \text{ steps}}
    + \eta \cdot \mathcal{L}_\pi, \qquad \eta = 0.10.
  \label{eq:joint}
\end{equation}
Gradients $\partial \mathcal{L}_\pi / \partial \phi$ flow back through
the forward pass that generated $\bfz_\mathrm{fused}$, causing the predictor
to favour representations that minimise future policy regret---not just
prediction accuracy. We call this the \emph{bidirectional gradient bridge}.

\subsection{Module 3: HierFed-KD}
\label{sec:hierfed}

\subsubsection{Zone FedAvg (Level 1)}
APs are grouped into $Z$ venue zones (departure terminals, arrival halls, etc.).
Every $T_\mathrm{local} = 100$ steps, quality-weighted FedAvg aggregates base
network parameters within each zone:
\begin{equation}
  \theta^\mathrm{zone}_z = \sum_{i \in \mathcal{Z}_z} w_i \theta_i,
  \quad w_i = \frac{\bar{r}_i}{\sum_j \bar{r}_j},
\end{equation}
where $\bar{r}_i$ is the exponential moving average reward of AP $i$.

\subsubsection{Global FedProx (Level 2)}
Every $T_\mathrm{global} = 500$ steps, zone-level models are aggregated globally
with a proximal penalty anchored at the \emph{previous} global model
$\theta_\mathrm{prev}$:
\begin{equation}
  \theta^* = \frac{\sum_z \theta^\mathrm{zone}_z + \mu \theta_\mathrm{prev}}{Z + \mu},
  \quad \mu = 0.01.
  \label{eq:fedprox}
\end{equation}
This differs from FedAvg in that it pulls the global model towards a stable
prior (the previous round), preventing drift from heterogeneous zone updates.

\subsubsection{Knowledge Distillation Compression}
Instead of transmitting raw weight tensors, each zone model $k$ shares
stochastic soft channel distributions $\pi_k^\mathrm{ch}$ (from Gumbel-softmax
with $\tau = 1.0$) and bandwidth predictions $\mathbf{w}_k$ evaluated over
$M = 100$ reference states. The global student model minimises:
\begin{equation}
  \mathcal{L}_\mathrm{KD} = \frac{1}{Z} \sum_k \left[
    \mathrm{KL}(\pi_k^\mathrm{ch} \| \hat{\pi}^\mathrm{ch})
    + \mathrm{MSE}(\mathbf{w}_k, \hat{\mathbf{w}})
  \right].
  \label{eq:kd}
\end{equation}
Soft distributions preserve policy diversity; sharing logits only (as in prior
work) would collapse diversity because deterministic queries yield one-hot outputs.
Communication savings: $M \times C \times 4 \text{ bytes}$ per zone $\ll$
full parameter count ($\sim 350\text{K}$ floats), yielding a $11.5\times$
compression ratio.

\subsubsection{Adaptive Lagrangian Multiplier}
AP capacity constraints are enforced without adding $r_\mathrm{cap}$ to the
environment reward (which would double-count the penalty). Instead, the
Lagrangian multiplier $\lambda_i$ updates exponentially:
\begin{equation}
  \lambda_i^{t+1} = \max\!\left(0, \lambda_i^t +
    \rho \cdot \max(0, L_i^t - 0.9)\right), \quad \rho = 0.01.
\end{equation}

The complete training procedure is described in Algorithm~\ref{alg:main}.

\begin{algorithm}[htbp]
\caption{STAG-FedSAC Training}\label{alg:main}
\begin{algorithmic}[1]
\State \textbf{Initialise} ST-GCAT $\phi$; actors $\pi_\theta^i$, critics $Q_\psi^i$,
       target critics $\bar{Q}^i$, replay buffers $\mathcal{B}_i$, Lagrangian
       $\lambda_i = 0$ for all $i$.
\For{episode $= 1, \dots, E$}
  \State $o \leftarrow \text{env.reset}()$
  \For{step $t = 1, \dots, T$}
    \State Build $\mathbf{H}, \mathbf{S}, \mathbf{A}$ from $o$ \Comment{History update: \emph{once} per step}
    \State $\bfz_\mathrm{fused}, \hat{L} \leftarrow \text{ST-GCAT}(\mathbf{H}, \mathbf{S}, \mathbf{A})$
    \For{$i = 1, \dots, N$}
      \State $a_i \sim \pi_\theta^i(\cdot | s_i)$ with $s_i$ from Eq.~\ref{eq:state}
    \EndFor
    \State $o', r, \mathrm{done} \leftarrow \text{env.step}(\{a_i\})$
    \For{$i = 1, \dots, N$}
      \State $\mathcal{B}_i.\mathrm{store}(s_i, a_i, r_i + \delta_p b_\mathrm{pred}
             + \lambda_i \cdot \text{cap\_violation}, s_i')$
    \EndFor
    \If{$t \ge T_\mathrm{warmup}$ \textbf{and} $t \bmod U_\mathrm{freq} = 0$}
      \State Update critics $Q^i$, actors $\pi^i$, temperature $\alpha$ via SAC
      \State \textbf{Joint step}: $\phi \leftarrow \phi - \eta_\phi
             \nabla_\phi \mathcal{L}_\mathrm{pred}$ (Eq.~\ref{eq:joint},
             using $\mathbf{H}$ read-only)
    \EndIf
    \State Update $\lambda_i$ for all $i$
    \If{$t \bmod T_\mathrm{local} = 0$} \textbf{Zone FedAvg} \EndIf
    \If{$t \bmod T_\mathrm{global} = 0$} \textbf{Global FedProx + KD} \EndIf
    \State $o \leftarrow o'$
  \EndFor
\EndFor
\end{algorithmic}
\end{algorithm}
