% ========================================================
% 5conclusion.tex
% ========================================================
\section{Conclusion}
\label{sec:conclusion}

We presented \textbf{STAG-FedSAC}, a unified framework that addresses four
open challenges in transit-venue WiFi management: schedule-blind prediction,
reactive control, unfair multi-class bandwidth allocation, and
communication-inefficient multi-AP coordination. The four novel contributions
---ST-GCAT with schedule cross-attention, graph-embedded SAC state,
bidirectional joint training, and HierFed-KD---were each demonstrated to be
independently necessary through a controlled ablation study. Together, they
achieve a collective $24.5\%$ reward improvement and $9.5\%$ fairness gain
over the strongest federated baseline on a 30-AP transit terminal simulation.

Three aspects of the framework deserve particular attention for future work:
\begin{enumerate}
  \item \textbf{NS-3 integration}: Replacing the Python simulator with an
  ns3-ai~\cite{ns3ai} interface would provide higher-fidelity SINR modelling,
  capturing hidden terminal effects and frequency-selective fading absent in
  the current Shannon approximation.

  \item \textbf{Real-world deployment}: The Dartmouth'18~\cite{DartmouthWiFi}
  and LCR HDD~\cite{LCRHDD} datasets confirm that real campus and stadium
  traces exhibit the surge patterns modelled here. A trial deployment at a
  high-speed rail station would validate transfer of the simulator-trained
  policies to hardware-level AP controllers.

  \item \textbf{Heterogeneous architectures}: As APs increasingly vary in
  radio chipset capability (Wi-Fi~6 vs.\ Wi-Fi~7), architecture-agnostic
  distillation beyond soft channel distributions---e.g., matching latent
  spaces via contrastive alignment~\cite{KDAFRL2025}---would strengthen
  HierFed-KD's applicability across heterogeneous venues.
\end{enumerate}

STAG-FedSAC's code and pre-trained checkpoints are publicly available at
\url{https://github.com/vinhqdang/STAG_FedSAC}, enabling reproducibility
and further community research on schedule-aware intelligent wireless systems.

\section*{Declaration of Competing Interest}
The authors declare no competing financial interests or personal relationships
that could have appeared to influence the work reported in this paper.

\section*{Data Availability}
The WACA Camp Nou dataset is publicly available at \url{https://doi.org/10.5281/zenodo.3960029}.
The Dartmouth WiFi traces are available at \url{https://doi.org/10.15783/C7F59T}.
Synthetic schedule data and simulation code are included in the public repository.

\section*{Acknowledgements}
The author thanks the British University Vietnam for providing access to the
NVIDIA RTX 5000 Ada GPU used for all experiments.
